MLH1 Results Outline

### 2020-03-02 doc/Results.rmd
getting hang of incorporating ref, use BibTex!

---
title: "MLH1 Results Outline"
date: '3/1/2020'
output:
  word_document:
    toc: yes
    toc_depth: '5'
  pdf_document: default
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
bibliography: refs_Main_Report.bib
---

@@

<!--
RW's main qtl paper used the within animal variance,

@article{wang2017genetics,
  title={Genetics of genome-wide recombination rate evolution in mice from an isolated Island},
  author={Wang, Richard J and Payseur, Bret A},
  journal={Genetics},
  volume={206},
  number={4},
  pages={1841--1852},
  year={2017},
  publisher={Genetics Soc America}
}
references for phenotypic variance connected to robustness to environment, increase in developmental consistency

@@

For each strain, we deconstructed the estimates of genome wide recombination rate by isolating single bivalents and quantifying the recombination landscape at the single bivalent level using an previously developed image analysis pipeline. The error of the image analysis algorithm is measuring chromosome features is low and highly similar to manual measures (Peterson 2019).

@@

For an additional  metric of chromatin compaction, we tested a whole cell measure of SC length using a second image analysis algorithm which converts the SC signal into single pixel wide trace of the SC shapes and provides accurate estimates of the summed SC length of bivalents for single cells (ref Wang). We apply the skeletonizing algorithm to all MLH1 meiocyte images. Mouse means were calculated from `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cell wide total SC measures (figure X). (where as `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` number of cells had MLH1 measures ).

@@

In all strains, the 1CO landscape had significant differences between males and female which followed the "typical landscape" of sex difference observed in many other species (Sardell Kirkpatrick). The foci position was more medially places in females while males have more terminal normalized foci positions (ttest; p = `r Q1_1CO.pos_t.test.FULL$p.value`).

@@

Crossover interference is one of the major determinates of the recombination landscape. It generates a distribution of evenly spaced crossover along chromosomes instead of a random and more uniform distribution. To quantifiy how crossover interference differs between the typical recombination landscapes of males and females, we isolated `r length(  (Curated_BivData.22 %>% filter(sex == "female") )$Obj.ID )` and 
`r length(  (Curated_BivData.22 %>% filter(sex == "male") )$Obj.ID )` 2CO bivalents in females and males respectively. Mean interfocal distances (IFD) were calculated from `r length( unique( (Curated_BivData.22 %>% filter(sex == "female") )$mouse) )`  female and `r length( unique( (Curated_BivData.22 %>% filter(sex == "male") )$mouse) )` male mice. We examined both raw IFD^raw^ and normalized by SC length (IFD^norm^). The mean IFD^raw^ measures of crossover interference as a mechanical force mediated through the SC. We compared the IFD^raw^ to test the hypothesis of interference acting as a mechanical force transmitted through the SC. Previous measures found no sex difference in measures of IFD^raw^ in mice (de Boer et al 2004).

>>>>

---
title: "MLH1 Results Outline"
date:
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
  word_document:
    toc: yes
    toc_depth: '5'
bibliography: refs_Main_Report.bib
---

@@

references for phenotypic variance connected to robustness to environment, increase in developmental consistency

@@

For each strain, we deconstructed the estimates of genome wide recombination rate by isolating single bivalents and quantifying the recombination landscape at the single bivalent level using an previously developed image analysis pipeline. The error of the image analysis algorithm is measuring chromosome features is low and highly similar to manual measures [@peterson_2019].

@@

For an additional  metric of chromatin compaction, we tested a whole cell measure of SC length using a second image analysis algorithm which converts the SC signal into single pixel wide trace of the SC shapes and provides accurate estimates of the summed SC length of bivalents for single cells [@wang_2019_sc]. We apply the skeletonizing algorithm to all MLH1 meiocyte images. Mouse means were calculated from `r length(  (MLH1.merge.Skel %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` cell wide total SC measures (figure X). (where as `r length(  (MLH1_data %>% filter(species == "M.musculus") %>% filter(subsp != "Cast")  )$Random.Name)` number of cells had MLH1 measures ).

@@

In all strains, the 1CO landscape had significant differences between males and female which followed the "typical landscape" of sex difference observed in many other species [@sardell_sex_2020]. The foci position was more medially places in females while males have more terminal normalized foci positions (ttest; p = `r Q1_1CO.pos_t.test.FULL$p.value`).

@@

Crossover interference is one of the major determinates of the recombination landscape. It generates a distribution of evenly spaced crossover along chromosomes instead of a random and more uniform distribution. To quantifiy how crossover interference differs between the typical recombination landscapes of males and females, we isolated `r length(  (Curated_BivData.22 %>% filter(sex == "female") )$Obj.ID )` and 
`r length(  (Curated_BivData.22 %>% filter(sex == "male") )$Obj.ID )` 2CO bivalents in females and males respectively. Mean interfocal distances (IFD) were calculated from `r length( unique( (Curated_BivData.22 %>% filter(sex == "female") )$mouse) )`  female and `r length( unique( (Curated_BivData.22 %>% filter(sex == "male") )$mouse) )` male mice. We examined both raw IFD^raw^ and normalized by SC length (IFD^norm^). The mean IFD^raw^ measures of crossover interference as a mechanical force mediated through the SC. We compared the IFD^raw^ to test the hypothesis of interference acting as a mechanical force transmitted through the SC. Previous measures found no sex difference in measures of IFD^raw^ in mice [@de_boer_two_2006].
